---
title: 'CMTH 642 Data Analytics: Advanced Methods Assignment 1'
output: pdf_document
Name: Amitkumar Singh
---

### 1. Read the csv files in the folder. (4 points)
```{r}
a<-read.csv("/Users/amit/Desktop/Big_data/CIND_642/USDA_Macronutrients.csv",header = TRUE)
b<-read.csv("/Users/amit/Desktop/Big_data/CIND_642/USDA_Micronutrients.csv",header = TRUE)

```

### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
USDA<-merge(a,b,by="ID")


View(USDA)
```

### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
```{r}
#str(USDA)
#sapply(USDA,class)
#sapply(USDA,typeof)
USDA$Sodium<-as.numeric(gsub(",","",USDA$Sodium))
USDA$Potassium<-as.numeric(gsub(",","",USDA$Potassium))

                                 

str(USDA)           


```

### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
USDAC <- USDA[rowSums(is.na(USDA))<=4,]
nrow(USDAC)

```

### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
USDA$Sugar[is.na(USDA$Sugar)] <- mean(USDA$Sugar, na.rm = TRUE)
USDA$VitaminE [is.na(USDA$VitaminE)]<-mean(USDA$VitaminE, na.rm = TRUE)
USDA$VitaminE [is.na(USDA$VitaminD)]<-mean(USDA$VitaminD, na.rm = TRUE)
View(USDA)
```

### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}

USDAclean<-USDA %>%drop_na()
any(is.na(USDAclean))
View(USDAclean)
```

### 7. Which food has the highest sodium level? (6 points)
```{r}
highsodium<-which.max(USDAclean$Sodium)
indexcol<- USDAclean[,2]
a<-max(USDAclean$Sodium)
a<-USDAclean[,8]
which.max(USDAclean[a])
names (USDAclean)
USDAclean[which.max(USDAclean$Sodium)]
```

### 8. Create a histogram of Vitamin C distribution in foods, with a limit of 0 to 100 on the x-axis and breaks of 100. (6 points)
```{r}
hist(USDAclean$VitaminC,xlim = c(0,100),breaks = 100)
```

### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
boxplot(USDAclean$TotalFat,USDAclean$Protein,USDAclean$Carbohydrate,main="TotalFat    Protien     Carbohydrate")
```

### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its calorie content. (6 points)
```{r}
plot(USDAclean$TotalFat,USDAclean$Calories,main = "Relationship Between Content & its Calorie")
```

### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise.Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
average<-USDAclean$Sodium
USDAclean$highlow
View(USDAclean)
apply
```

### 12. Calculate the average amount of iron, sorted by high and low protein. (8 points)
```{r}
USDAclean$Iron
```

### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
install.packages("jpeg")
library(jpeg)
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
# INSERT YOUR ANSWER HERE
```

### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
# INSERT YOUR ANSWER HERE
```

### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
# INSERT YOUR ANSWER HERE
```

### 16. Save your final data frame as "USDAclean_ [your last name]" (4 points)
```{r}
# INSERT YOUR ANSWER HERE
```


